---
title: "Untitled"
author: "Savannah Mierau"
date: "9/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


```{r}
library(knitr)
library(tidyverse)

leadiq <- read.csv("lead-iq-01.csv")
# A graph showing the IQ levels by location status
qplot(Smelter, IQ, data=leadiq)

#At least one nicely formatted table using the 'kable' function,
#This table shows each column in the data set
kable(
  leadiq,
  format = "html")

#This table gives a summary of data looking at the amount of observations for each variable and the classification of the variables (character in this situation)
kable(t(apply(leadiq,2,summary)))

#Inline calculations giving the values of the means, that will be updated if the data are changed
mean(leadiq$IQ)
sd(leadiq$IQ)

```
A couple sentences of text (not on the graph, but in the body of the Rmarkdown document) describing the graph and the table
Graph
Figure 1 shows the varying IQ levels in children with high lead levels or low lead levels determined by distance from the smelter. IQ levels among the Far group tend to be higher, but there is an outlier that is likely pulling the data for the Far group. The data should be investigated after removing the outlier. 
Table
Table 1 shows IQ levels based on location from the smelter. 

3. There is an id that is pulling the "Far" data. 

```{r}
# removing outlier
omit <- c("999")
# making new dataset without the outlier
leadiq <- leadiq[!leadiq$IQ %in% omit,]

mean(leadiq$IQ)
sd(leadiq$IQ)

```



